Predicting Remaining Useful Life of NASA Aircraft Engines

### 2025-02-25 index.qmd
added label for test/train
Updated references
Added Meier Kaplan

Each engine has an unknown amount of wear, manufacturing variation, and
sensor noise all of which impact the survival time. In the training
dataset, each engine is operating normally at the beginning of the time
series, develops a fault, and experiences failure at the end of the time
series. The conditions the engines are subjected to are represented by
three operational setting fields and 21 sensor measurement fields. In
the testing data, the engines have not experienced failure
[@NASA_CMAPSS].

```{r, echo = FALSE}
#packages
library(tidyverse)
library(gridExtra)
library(survival)
library(survminer)

@@

@@

test <- read.table(testurl, header = TRUE, sep = '') %>%
  mutate(status = 0)
head(test, n = 5)
```

@@

train <- read.table(trainurl, header = TRUE, sep = '') %>%
  group_by(id) %>%
  mutate(status = if_else(row_number() == n(), 1, 0)) %>%
  ungroup()

@@

@@

univ_models <- lapply( univ_formulas, function(x){coxph(x, data = train)})
# Extract data
univ_results <- lapply(univ_models,
                       function(x){
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coeficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (",
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test",
                                        "p.value")
                          return(res)
                         })
as.data.frame(univ_results)
```

>>>>

Each engine has an unknown amount of wear, manufacturing variation, and
sensor noise all of which impact the survival time. In the training
dataset, each engine is operating normally at the beginning of the time
series, develops a fault, and experiences failure at the end of the time
series. The conditions the engines are subjected to are represented by
three operational setting fields and 21 sensor measurement fields. In
the testing data, the engines have not experienced failure
[@saxena2008].

```{r, echo = FALSE}
#packages
library(tidyverse)
library(gridExtra)
library(survival)
library(survminer)
library(ggpubr)
library(magrittr)
library(ggfortify)

@@

**Testing Data**

@@

test <- read.delim(testurl, header = TRUE, sep = '') %>%
  mutate(status = 0)
head(test, n = 5)
```

**Training Data**

@@

train <- read.delim(trainurl, header = TRUE, sep = '') %>%
  group_by(id) %>%
  mutate(status = if_else(row_number() == n(), 1, 0)) %>%
  ungroup()

@@

The Kaplan-Meier survival curve is shown below.

```{r}
t <- as.matrix(train[, "time"])
s <- as.matrix(train[, "status"])

km_fit <- survfit(Surv(t, s) ~ 1)

autoplot(km_fit) + theme_bw()
```

@@

univ_models <- lapply( univ_formulas, function(x){coxph(x, data = train)})

# Extract data
univ_results <- lapply(univ_models,
                       function(x){
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coefficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (",
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test",
                                        "p.value")
                          return(res)
                         })
as.data.frame(univ_results)
```
